### 2025-11-01 thesis/implementation/index.tex
Embedding and masking section

>>>>

\subsubsection{Embedding and Masking}

The embedding and masking functionalities in \texttt{natural-sampling} abstract
batch embedding of masked question and SQL queries using the GPU so that the
runtime and sampling code can efficiently compute masked embeddings.

\paragraph{Architecture}

The two primary types for embedding plain and masked text are the
\texttt{SemanticString} and \texttt{MaskedSemanticString} structs which
encapsulate strings alognside their embedding vectors. The
\texttt{MaskedSemanticString} version additionally contains the masked and
unmasked versions of the embedded string. Furthermore they offer
\texttt{embed\_seq} and \texttt{embed\_chunked} methods which can be used for
batch computation of embeddings.

\paragraph{Batch Computation}

During sampling, offloading data to the GPUs memory is an expensive operation.
When sampling hundred thousands or millions of samples, incurring the IO
roundtrip from CPU to GPU for computing a single embedding is a
performance bottleneck that increases the sampling time to an order of days
(eg, when sampling the \textsc{SynSQL}) dataset. Thus batch processing can help to
minimise the IO roundtrips needed between the CPU and GPU before computing
embeddings and better utilise the available computing power.

Using the \texttt{embed\_seq} method, a sequence of strings can be embedded and
a sequence of \texttt{SemanticStrings} is returned. Furthermore through the
\texttt{embed\_chunked} method, chunks of unrelated data can be embedded in a
single operation. The method maintains the original chunk layout which makes it
possible to reconstruct the input semantic seamlessly. This optimization is
used to embed the questions and sql queries alongside while maintaining clear
separation in the embedding output.

\begin{verbatim}
    INSERT DIAGRAM
\end{verbatim}

% NOTE: Batch embedding computation time down from 200ms per item to
% approximately 2ms per item when processing 1000-item batches.
% PUT INTO TEXT/diagram description

\paragraph{Integration}

The \texttt{masking} module tightly integrates wiuth the embedding pipeline.
During sampling, questions and queries are first masked, then embedded together
using \texttt{embed\_chunked} and finally combined into
\texttt{MaskedSemanticString} instances for storage in \texttt{Vector}.

\subsubsection{Wasserstein-Weisfeiler-Leman Kernels}

\subsection{Benchmarking Infrastructure}

\textsc{Natural}'s benchmarking infrastructure is implemented in
\texttt{natural-benchmark}. This benchmarking infrastructure is enabling the
development, verification and ultimately deployment of \textsc{Natural}
pipeline. This section focuses on the respective benchmarking infrastructure
development, performance optimizations that were required and engineering
challenges encountered.

Whilst the benchmarking infrastructure is not part of the core pipeline, they
help to develop and test hypotheses for extensions or design changes
confidently, compare performance across pipeline versions, models and
understand the benchmarking data.
